\documentclass{article}
\usepackage[en]{homework}

\config{Algebra\ -\ 0}{2026 Spring}{1}

\begin{document}

\begin{problem}[1B-2]
Suppose $ a\in\mathbb{F} $, $ v\in V $, and $ av=0 $. Prove that: $ a=0 $ or $ v=0 $.
\end{problem}

\begin{proof}
Assume that $ a\neq 0 $, then it's sufficient to prove $ v=0 $.\par 
By the associativity of the scalar multiplication, $ v=(a^{-1}a)v=a^{-1}(av)=a^{-1}0 $. By the property of the zero vector, $ a^{-1}0=0 $. Thus $ v=0 $.
\end{proof}

\begin{problem}[1B-3]
Suppose $v, w \in V$. Explain why there exists a unique $x \in V$ such that $v + 3x = w$.
\end{problem}

\begin{solution}
The addictive inverse of $ v $ is unique, denoted as $ -v $. Then $ x=\frac{1}{3}(w-v) $ is the unique solution to the equation $ v+3x=w $.
\end{solution}

\begin{problem}[1B-5]
Show that in the definition of a vector space (1.20), the additive inverse condition can be replaced with the condition that
\[ 0v=0 \text{ for all }v\in V. \]
Here the 0 on the left side is the number 0, and the 0 on the right side is the additive identity of $ V $.
\end{problem}

\begin{proof}
For every $ v\in V $, define the addictive inverse of $ v $ is $ (-1)v $. Then $ v+(-1)v=v+(-1)1v=(1-1)v=0v=0 $. Thus the addictive inverse condition is satisfied.
\end{proof}

\begin{problem}[1B-6]
Let $\infty$ and $-\infty$ denote two distinct objects, neither of which is in $\R$. Define an addition and scalar multiplication on $\R \cup \{ \infty, -\infty \}$ as you could guess from the notation. Specifically, the sum and product of two real numbers is as usual, and for $t \in \R$ define
\[
t\infty = 
\begin{cases} 
-\infty & \text{if } t < 0, \\
0 & \text{if } t = 0, \\
\infty & \text{if } t > 0.
\end{cases}
\quad
t(-\infty) = 
\begin{cases} 
\infty & \text{if } t < 0, \\
0 & \text{if } t = 0, \\
-\infty & \text{if } t > 0.
\end{cases}
\]
and
\[
t + \infty = \infty + t = \infty + \infty = \infty,
\]
\[
t + (-\infty) = (-\infty) + t = (-\infty) + (-\infty) = -\infty,
\]
\[
\infty + (-\infty) = (-\infty) + \infty = 0.
\]
With these operations of addition and scalar multiplication, is $\R \cup \{ \infty, -\infty \}$ a vector space over $\R$? Explain.
\end{problem}

\begin{solution}
No. The distributive law fails, since
\[ (-\infty+\infty)+\infty= 0+\infty = \infty,\text{ but } -\infty+(\infty+\infty) = -\infty+\infty = 0. \]
\end{solution}

\begin{problem}[1C-1]
For each of the following subsets of $\mathbb{F}^3$, determine whether it is a subspace of $\mathbb{F}^3$.

\begin{enumerate}[label=(\alph*)]
    \item $\{ (x_1, x_2, x_3) \in \mathbb{F}^3 : x_1 + 2x_2 + 3x_3 = 0 \}$
    \item $\{ (x_1, x_2, x_3) \in \mathbb{F}^3 : x_1 + 2x_2 + 3x_3 = 4 \}$
    \item $\{ (x_1, x_2, x_3) \in \mathbb{F}^3 : x_1 x_2 x_3 = 0 \}$
    \item $\{ (x_1, x_2, x_3) \in \mathbb{F}^3 : x_1 = 5x_3 \}$
\end{enumerate}
\end{problem}

\begin{solution}
\begin{enumerate}[label=(\alph*)]
    \item Yes. The zero vector is in the set, and the set is closed under addition and scalar multiplication.
    \item No. The zero vector is not in the set.
    \item No. The set is not closed under addition, since $ (1, 0, 0) $ and $ (0, 1, 1) $ are in the set but their sum $ (1, 1, 1) $ is not in the set.
    \item Yes. The reason is similar to (a).
\end{enumerate}
\end{solution}

\begin{problem}[1C-5]
Is $ \R^2 $ is a subspace of the complex space $ \C^2 $?
\end{problem}

\begin{solution}
No, since \(\R^2\) is not closed under scalar multiplication with respect to an imaginary scalar in \(\C\).
\end{solution}

\begin{problem}[1C-6]
\begin{enumerate}[label=(\alph*)]
    \item Is $\{(a, b, c) \in \R^3 : a^3 = b^3\}$ a subspace of $\R^3$?
    \item Is $\{(a, b, c) \in \C^3 : a^3 = b^3\}$ a subspace of $\C^3$?
\end{enumerate}
\end{problem}

\begin{solution}
\begin{enumerate}[label=(\alph*)]
    \item Yes. $ a^3=b^3, a,b\in\R $ implies $ a=b $, then the set is equal to $\{(a, b, c) \in \R^3 : a = b\}$, which is a subspace of $\R^3$.
    \item No. The set is not closed under addition, since $ (1, 1, 0) $ and $ (1, \omega, 0) $ are in the set but their sum $ (2, 1+\omega, 0) $ is not in the set, where $ \omega=(-1+\sqrt{3}\i)/2 $.
\end{enumerate}
\end{solution}

\begin{problem}[1C-12]
Prove that the union of two subspaces of $ V $ is a subspace of $ V $ if and only if one of the subspaces is contained in the other.
\end{problem}

\begin{proof}
The sufficiency is trivial, so we will only prove the necessity.\par 
Conversly, assume that $ U_1, U_2 $ and $ U:=U_1\cup U_2 $ are subspaces of $ V $, but neither $ U_1\subset U_2 $ nor $ U_2\subset U_1 $. Then, there exist $ u_1\in U_1 - U_2 $ and $ u_2\in U_2 - U_1 $.\par
Since $ U $ is closed under addition, $ u_1+u_2\in U $, which implise ether $ u_1+u_2\in U_1 $ or $ u_1+u_2\in U_2 $. However, if $ u_1+u_2\in U_1 $, then $ u_2=(u_1+u_2)-u_1\in U_1 $, which contradicts the choice of $ u_2 $. Similarly, if $ u_1+u_2\in U_2 $, then $ u_1\in U_2 $, which contradicts the choice of $ u_1 $. Thus we get a contradiction, and the necessity is proved. 
\end{proof}

\begin{problem}[1C-13]
$ V $ is a vector space over $ \F $, where $ \F\neq \F_2 $ . Prove that the union of three subspaces of $ V $ is a subspace of $ V $ if and only if one of the subspaces contains the other two.
\end{problem}

\begin{proof}
The sufficiency is trivial, so we will only prove the necessity.\par
Assume there are three subspaces $ U_1, U_2, U_3 $ of $ V $ such that $ U:=U_1\cup U_2\cup U_3 $ is a subspace of $ V $. If $ U_2\subset U_1 $ or $ U_1\subset U_2 $ (WLOG the former), then $ U=U_1\cup U_2\cup U_3=U_1\cup U_3 $ . By the previous problem, either $ U_1\subset U_3 $ or $ U_3\subset U_1 $, and both cases meet the requirement.\par
Otherwise, if  $ U_1\not\subset U_2 $ and $ U_2\not\subset U_1$ , then we're going to prove: $ U_1\subset U_3 $ and $ U_2\subset U_3 $.\par 
For every $ u_1\in U_1-U_2 $ and $ u_2\in U_2-U_1 $ , $ u_1+u_2\in U $ . If $ u_1+u_2\in U_1 $, then $ u_2=(u_1+u_2)-u_1\in U_1 $, which contradicts the choice of $ u_2 $. Similarly, if $ u_1+u_2\in U_2 $, then $ u_1\in U_2 $, which contradicts the choice of $ u_1 $. Thus $ u_1+u_2\in U_3 $. \par 
 $ \F\neq\F_2 $ , thus $ 2u_1\neq 0 $ , then $ 2u_1\in U_1-U_2 $ . For the same reason,  $ 2u_1+u_2\in U_3 $ . Hence  $ u_1=(2u_1+u_2)-(u_1+u_2)\in U_3 $ . Hence we have proved  $ U_1-U_2\subset U_3 $ .\par 
So, to prove  $ U_1\subset U_3 $ , it's sufficient to prove  $ U_1\cap U_2\subset U_3 $. Note that $ W:=U_1\cap U_2 $ is a subspace of  $ U_1 $ ( $ W $ is closed under addition and scalar multiplication). Choose  $ v\in U_1-W=U_1-U_2\subset U_3$, then for every  $ w $ in $ W $ , $ v+w\not\in W $(otherwise $ v=(v+w)-w\in W $), then $ v+w\in U_1-U_2\subset U_3 $. Hence  $ w=(v+w)-v\in U_3 $. Thus $ W\subset U_3 $. \par 
That's what we want to prove, and the necessity is proved.
\end{proof}

\begin{problem}[1C-21]
Suppose
\[
U = \{(x, y, x + y, x - y, 2x) \in \mathbb{F}^5 : x, y \in \mathbb{F}\}.
\]
Find a subspace \( W \) of \( \mathbb{F}^5 \) such that \( \mathbb{F}^5 = U \oplus W \).
\end{problem}

\begin{solution}
Let  $ W=\{(0,0,a,b,c)\in\F^5: a,b,c\in\F\} $. We'll prove this $ W $ meets the requirement.\par 
First, we're going to prove: $ U+V $ is direct sum . This is equivalent to prove: $ (x,y,x+y,x-y,2x)+(0,0,a,b,c)=0 $ implies  $ x=y=a=b=c=0 $ . This is obvious, since the equation w.r.t the first two components implies $ x=y=0 $, then follows $ a=b=c=0 $. Thus  $ U+V $ is direct sum .\par
Then to prove  $ U\oplus W=\F^5 $ . That's clear, since
\[ (x_1,x_2,x_3,x_4,x_5)=(x_1,x_2,x_1+x_2,x_1-x_2,2x_1)+(0,0,x_3-x_1-x_2,x_4-x_1+x_2,x_5-2x_2). \] 
\end{solution}

\begin{problem}[1C-24]
A function \( f : \mathbb{R} \to \mathbb{R} \) is called \emph{even} if
\[
f(-x) = f(x)
\]
for all \( x \in \mathbb{R} \). A function \( f : \mathbb{R} \to \mathbb{R} \) is called \emph{odd} if
\[
f(-x) = -f(x)
\]
for all \( x \in \mathbb{R} \). Let \( V_\mathrm{e} \) denote the set of real-valued even functions on \( \mathbb{R} \) and let \( V_\mathrm{o} \) denote the set of real-valued odd functions on \( \mathbb{R} \). Show that
\[
\mathbb{R}^{\mathbb{R}} = V_\mathrm{e} \oplus V_\mathrm{o}.
\]
\end{problem}

\begin{proof}
First, we're going to prove: $ V_\mathrm{e}+V_\mathrm{o} $ is direct sum . This is equivalent to prove  $ V_\mathrm{e}\cap V_\mathrm{o}=\{0\} $(Notice that there're only two subspaces to be added). For  $ f\in V_\mathrm{e}\cap V_\mathrm{o}$, we have  $ f(x)=f(-x)=-f(x) $ for every $ x\in\R $, then $ f=0 $. Thus $ V_\mathrm{e}+V_\mathrm{o} $ is direct sum .\par
Then to prove  $ V_\mathrm{e}\oplus V_\mathrm{o}=\R^\R $ . For every $ f\in\R^\R $, define $ f_\mathrm{e}(x)=(f(x)+f(-x))/2 $ and $ f_\mathrm{o}(x)=(f(x)-f(-x))/2 $. Then it's easy to verify that  $ f=f_\mathrm{e}+f_\mathrm{o} $, and  $ f_\mathrm{e}\in V_\mathrm{e} $,  $ f_\mathrm{o}\in V_\mathrm{o} $. Thus  $ V_\mathrm{e}\oplus V_\mathrm{o}=\R^\R $. That's what we want to prove, and the problem is solved.   
\end{proof}


\end{document}